---
title: "Recommender System"
author: "Anthony A"
date: "9/19/2022"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
library(magrittr)
library(ggplot2)
library(dplyr)
library(purrr)
library(testthat)
```

# Abstract

A recommender system is intended to predict the sentiment of users based on the collective population.
In the movie streaming services such as Amazon, we can generate user recommendations for movies.
The idea is that watchers conform to a niche so they will continue to watch and rate those movies highly.
By finding out what half of the group liked to watch, we can predict that the other half will also be interested.
Thus, when collective users of the same niche exist, these groups of data observations are called clusters.

```{r initializes}
### Shortens column names for printing
ShortenTitles <- function(df, n = 12) {
  magrittr::set_colnames(df, substring(colnames(df), 1, n))
}

### Generates an n-length vector of integers 1 through 5
generatedVector = set.seed(8675309) %>%
  { rnorm(240, 3, 2) } %>%
    { . %% 5 } %>%
      ceiling(.) %>%
        { (function(x) {
          deleteVals = runif(length(x)) > 0.5
          x[deleteVals] = NA
          return(x)
        })(.) }

### Visualizes the distribution of randomized data in a histogram
generatedVector %>%
  .[! is.na(.)] %>%
    data.frame(n = .) %>%
      ggplot2::ggplot(.) +
      ggplot2::aes(x = n) +
      ggplot2::geom_histogram(bins = 5L, fill = "#112446", col="grey") +
      ggplot2::geom_density(alpha = 0.6, fill = "red")
```

# Datasets

Data generation is used in lieu of actual data.
The data has to be pruned of empty rows or else there is no baseline to estimate the customer's desires.
As a result, a recommender system has to recommend the top 10 movies of all time for a brand new customer.

```{r data, warning = FALSE}
### Top ten movies of 1963
movieTitles = c(
  "Cleopatra", "How the West Was Won", "It's a Mad, Mad, Mad, Mad World",
  "Tom Jones", "Irma la Douce", "Son of Flubber", "Charade", "Bye Bye Birdie",
  "Come Blow Your Horn", "Move Over, Darling", "The Great Escape"
)

### Converts generated data into a matrix and removes empty rows
data = as.data.frame(matrix(generatedVector, ncol = 6)) %>%
  magrittr::set_colnames(., movieTitles[1:ncol(.)]) %>%
    .[apply(., 1, function(x) { any(! is.na(x))}), ]

### Check data for cleanliness on a pre-written capacity
testthat::test_that("All rows have at least one value", {
  testthat::expect_true(all(apply(data, 1, function(x) {
    any(! is.na(x))
  })))
})
```

# Global Baselines

There are three values being used to assess the recommended movie for a customer.
The user baseline and movie baseline are the average ratings of each user and movie.
Afterwards, we average every seen observation for the overall movie baseline.
These are added together as such: Users + Movie - 2 * Average

```{r globalBaselines}
### Calculates the mean of each row and column
userBaselines = apply(data, 1, mean, na.rm = TRUE)
movieBaselines = apply(data, 2, mean, na.rm = TRUE)

### Calculates the mean of every available operation
avgMovieRating = mean(unlist(data), na.rm = TRUE)


movieBaselines
```

# Known observations

The input data shown below is a glimpse of what data is available to our recommender system.
Each row is a unique watcher.
These are the ratings given by users who have already watched the movies.

```{r givenRatings}
### Show the input data
head(ShortenTitles(data))
```

# Predicted ratings

After running our algorithm on all of the movies, we get a new data set where every movie has a rating.
Generally, the expectation is that certain movies will cluster and form visible trends.
These trends will be picked up and inquiry to the system would return the highest value unwatched movies.
Our glimpse at the top shows fairly accurate results on the right side but completely erroneous ones on the left side.

```{r artificialRatings}
### Generates predictions from available data
artificialRatings = matrix(-avgMovieRating, nrow = length(userBaselines), ncol = length(movieBaselines)) %>%
  { . + movieBaselines } %>%
    t(.) %>%
      { . + userBaselines } %>%
        t(.)

### Show the created predictions
data.frame(artificialRatings) %>%
  magrittr::set_colnames(., movieTitles[1:ncol(.)]) %>%
    ShortenTitles(.) %>%
      format(., digits = 3) %>%
        head(.)
```

# Model Verification

It is important to verify that the predictions match the reality of the situation.
By comparing the real data to the predicted values, we should run a statistical analysis.
The mean and and median for this particular set makes it seem like it worked.
However, the two extremes are +/- 4 which would be the exact *opposite* of the intended rating.

```{r verification}
### Generate data to see where historical data differs from predictions
confusionMatrix = data %>%
  { (function(df) {
    x = is.na(df)
    df[! x] = df[! x] - artificialRatings[! x]
    df
  })(.) }

### Get summary data to understand model accuracy
confusionMatrix[! is.na(confusionMatrix)] %>%
  unlist(.) %>%
    summary(.)
```

This problem is even more apparent when the data is graphed.
There is a bell shape because the randomly generated data is normally distributed.
However, about half of the predictions are incorrect and too far away from zero.
Hypothetically, a real dataset would exhibit clustering and be largely close to zero.
Since the data is random with no favored clusters, the result is also random as shown below.

```{r verificationGraph}
### Generates a visualization of model accuracy
confusionMatrix %>%
  unlist(.) %>%
    .[! is.na(.)] %>%
      data.frame(n = .) %>%
        ggplot2::ggplot(.) +
        ggplot2::aes(x = n) +
        ggplot2::geom_density(adjust = 0.2, fill = "#112446") +
        ggplot2::theme_bw()
```

