---
title: "Using NY Times Web API"
author: "Anthony A"
date: "11/30/2022"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
library(magrittr)
library(httr)
library(jsonlite)
```

# Abstract

Using the NY Times API, we can retrieve data that corresponds to our needs.
Instead of manual web scrapping or writing emails, we can programmatically assess bestselling books.
This is done with an HTML request which returns a JSON data structure.
It is a highly efficient work practice that is used everyday because it makes open data accessible.

```{r initializes, include=FALSE}
apiKey = "Y9o5BDG2LUOIJGGzV6Z403e73xkqymzU"
```

# Data Sources

The NY Times API is documented and can be found here:
https://developer.nytimes.com/docs/books-product/1/overview

```{r constructors}
ConstructApiCall <- function(url, ..., key = paste0("api-key=", apiKey)) {
  paste0(url, "?", paste(..., key, sep = "&"))
}

ApiCall <- function(url) {
  results = httr::GET(url)
  results$url = gsub("api-key=.*", "", results$url)
  results$request$url = gsub("api-key=.*", "", results$request$url)
  
  if (results$status_code >= 100 & results$status_code <= 299) {
    return(rawToChar(results$content))
  } else {
    warning(results$status_code)
    return(rawToChar(results$content))
  }
}

LessWords <- function(df) {
  apply(df, 2, substring, 0L, 20L)
}
```

# Available Datasets

According to the NY Times Books API, they divide book into categories and the list is pulled through the API.
Thus, we can use the GET /lists/names.json endpoint to get the bestsellers types of books.
This will help us refine our search later to non-fiction.

```{r data, warning = FALSE}
timesBaseUrl = "https://api.nytimes.com/svc/books/v3/"

paste0(timesBaseUrl, "/lists/names.json") %>%
  ConstructApiCall(.) %>%
    ApiCall(.) %>%
      jsonlite::fromJSON(.) %>%
        .$results %>% 
          .[grepl("nonfiction", .[[3]], ignore.case = TRUE), ]
```

# Non-Fiction Hardcover Books

Now that we know the necessary encoded name, we can construct a call on the API.
With the GET /lists.json endpoint, we can get a vector and convert it to a data frame.
The parameters are appended to the end of the URL to specify list and date.
These are called query parameters.

```{r verification}
paste0(timesBaseUrl, "lists.json") %>%
  ConstructApiCall(., "list=hardcover-nonfiction", "published-date=2014-04-20") %>%
    ApiCall(.) %>%
      jsonlite::fromJSON(.) %>%
        .$results %>%
          .$book_details %>%
            do.call(rbind, .) %>%
              LessWords(.) %>%
                head(.)
```

# Ranking the Non-Fiction Hardcover Books

Alternatively, they provide another endpoint where the book details are not nested.
This is preferable for our use case if we can start with a data frame instead of making it ourselves.
The GET /lists/{date}/{list}.json endpoint will give us all non-fiction books from 2014.
These correspond to path parameters because they are embedded in the URL.
Path parameters are adjusted by inserting them directly into the link as shown.
If we had more than 20 results, an optional query parameter is provided as the offset.

```{r}
paste0(timesBaseUrl, "lists/2014-04-20/hardcover-nonfiction.json") %>%
  ConstructApiCall(., "offset=0") %>%
    ApiCall(.) %>%
      jsonlite::fromJSON(.) %>%
        .$results %>%
          .$books %>%
            LessWords(.) %>%
              head(.)
```

